---
title: "School Ranking Vizualization"
author: "Emilia Pawlowska"
date: "2024-01-15"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For the final assignment I chose topic "3. Analyzing number of schools".
The aim of the project is to analyze the number and location of high schools in Gdynia along with information about the year of their establishment and their place in the "Pomorski Ranking Liceów Ogólnokształcących 2023" prepared by the portal Perspektywy. We want to create a leaflet map and show schools as a point with pop up menu where the marker color corresponds to their rank in the ranking (gold, silver or bronze). 

The method of visualizing high schools in Gdynia on a map, as presented in the project, can be very valuable for primary school students seeking information about prospective schools. 

Ranking site: https://2023.licea.perspektywy.pl/2023/rankings/ranking-pomorski

Other information were collected from schools home pages.

<!--Installing libraries-->

```{r libraries, include=FALSE}
#install.packages("leaflet")
library(leaflet)
```

# Creating data for analyzing high schools in Gdynia
- latitude and longitude of the school
- year of establishing
- place in the ranking
- rank (Quality mark in Perspektywy Ranking)

In case of the vector with ranks I decided to use numerical value where: 1 - gold, 2 -silver, 3- bronze.
```{r}
schools_names <- c("III Liceum Ogólnokształcące im. Marynarki Wojennej RP", 
                   "VI LO im. Wacława Sierpińskiego", 
                   "X Liceum Ogólnokształcące", 
                   "Katolickie Liceum Ogólnokształcące im. Św. Jana Pawła II",
                   "II LO im. Adama Mickiewicza", 
                   "XIV LO z Oddziałami Dwujęzycznymi im Mikołaja Kopernika", 
                   "I LO im. Zasłużonych Ludzi Morza",
                   "IX LO im. Marszałka Józefa Piłsudskiego",
                   "XVII LO",
                   "Ogólnokształcące Liceum Jezuitów Z Oddziałami Dwujęzycznymi Im Św. Stanisława Kostki")

ranking_places <- c(1, 7, 12, 17 ,10, 23, 37, 35, 24, 36)

year_of_establishment  <- c(1950, 1972, 1953, 1992, 1931, 1993, 1948, 1991, 2019, 1937)

ranks <- c(1, 1, 1, 2, 1, 2, 3, 3, 2, 3)

longitudes <-c(18.544909252828344,
               18.540935474697612,
               18.53672274788308,
               18.509271985362744,
               18.519644278602872,
               18.466448898992436,
               18.51625368474876,
               18.55039783955798,
               18.428238980864435,
               18.52564368421164)

latitudes <-c(54.507172618186885,
             54.50304147439088,
             54.51109308175149,
             54.55285195157312,
             54.519311780444525,
             54.54236267453898,
             54.49857389963467,
             54.47838490037935,
             54.48848516237254,
             54.51566908702809)

```


### Creating a dataframe

Creating a dataframe with 6 columns: School, Latitude, Latitude, Year, Place, Rank.
```{r}
schools_df <- data.frame(School = schools_names,
                         Latitude = latitudes,
                         Longitude = longitudes,
                         Year = year_of_establishment,
                         Place = ranking_places,
                         Rank = ranks)
```


# Map of schools in Gdynia
## Creating custom icons and a map using Leaflet

Firstly, I found the coordinates of Gdynia for setView() so that the map view would be set in the right place. I decided to use addAwesomeMarkers to mark schools as points because it allows me to customize icons. For customization, I created a getColor function, which assigns a color to the marker depending on the school's rank. I also added a legend for the map's readability.
Thanks to the popup, after clicking on the marker information about name, city, year, and place in the ranking is displayed. The value for City is static because the project includes only schools from Gdynia. 

```{r, echo=FALSE}

getColor <- function(schools_df) {
  sapply(schools_df$Rank, function(Rank) {
  if(Rank <= 1) {
    "orange"
  } else if(Rank == 2) {
    "lightgray"
  } else {
    "darkred"
  } })
}

gdynia_coordinates <- c(54.5, 18.43)


leaflet() %>%
  setView(lng = gdynia_coordinates[2], lat = gdynia_coordinates[1], zoom = 11) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addAwesomeMarkers(
    data = schools_df,
    lng = schools_df$Longitude,
    lat = schools_df$Latitude,
    icon=awesomeIcons(
        markerColor = getColor(schools_df)
    ),
    popup = paste0(
      "<b>School name: </b>", schools_df$School, "<br>",
      "<b>City: </b>", "Gdynia" ,"<br>",
      "<b>Year: </b>", schools_df$Year, "<br>",
      "<b>Ranking place: </b>", schools_df$Place,"<br>"
    )
  ) %>%
   addLegend(
    position = "bottomright",
    colors = c("orange", "lightgray", "darkred"),
    labels = c("Gold", "Silver", "Bronze"),
    title = "\"Perspektywy\" Ranking"
  )

```







